---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geometries

<!-- badges: start -->
<!-- badges: end -->

Have you every wanted to generate geometric structures from data.frames, but independant of any R classes, attributes or libraries? 

No? Ok, this library isn't for you.

But if you answered 'yes', this might be of interest.

## Shapes

When one thinks of data structures of 'shapes' in R, the current standard / typical structures are

- **Point** - a vector of x, y values
- **Line** - a matrix of x, y columns
- **Polygon** - a list of matrices of x, y columns
- **MultiPolygon** - a list of list of matrices of x, y columns

(in reality you can have more than just x & y columns)

But constructing these can sometimes be a bit fiddly. 

So you can use the **shapes** API to build these structures from tabular data (e.g. data.frames)

```{r}

df <- data.frame(
  x = 1:10
  , y = 10:1
  , z = 21:30
  , m = 30:21
  , val = letters[1:10]
  , id = c( rep(1,5), rep(2,5) )
)

df

library(Rcpp)

cppFunction(
  depends = "geometries"
  , includes = '#include "geometries/shapes/shapes.hpp"'
  , code = '
    SEXP line( SEXP df, SEXP geometry_cols ) {
      return geometries::shapes::get_mat( df, geometry_cols );
    }
  '
)

line( df, c("x","y") )


cppFunction(
  depends = "geometries"
  , includes = '#include "geometries/shapes/shapes.hpp"'
  , code = '
    SEXP polygon( SEXP df, SEXP geometry_cols, SEXP line_id ) {
      return geometries::shapes::get_listMat( df, geometry_cols, line_id );
    }
  '
)

polygon( df, c("x","y"), c("id") )

cppFunction(
  depends = "geometries"
  , includes = '#include "geometries/shapes/shapes.hpp"'
  , code = '
    SEXP multipolygon( SEXP df, SEXP geometry_cols, SEXP polygon_id, SEXP line_id ) {
      return geometries::shapes::get_listListMat( df, geometry_cols, polygon_id, line_id );
    }
  '
)

multipolygon( df, c("x","y"), c("id"),  c("id") )


```


## Collections


## Interleaving



## Utilities

```{r}

library(Rcpp)

## calculate_bbox
## returns the xmin, ymin, xmax and ymax of various objects

cppFunction(
  depends = "geometries"
  , includes = '#include "geometries/bbox/bbox.hpp"'
  , code = '
    Rcpp::NumericVector calculate_bbox( SEXP x, SEXP geometry ) {
      return geometries::bbox::calculate_bbox( x, geometry );
    }
  '
)

## of a data.frame
df <- data.frame(
  x = 1:10
  , y = 21:30
  , z = 31:40
)

calculate_bbox(df, c("x","y"))
calculate_bbox(df, c("x","z"))
calculate_bbox(df, c("z","y"))

## If using a list, 
l <- list(
  matrix(1:6, ncol = 2 )
  , matrix(6:1, ncol = 2)
)

calculate_bbox( l, NULL )

l <- list(
  matrix(1:6, ncol = 3 )
  , matrix(6:1, ncol = 3)
)

calculate_bbox( l, c(0,1) )
calculate_bbox( l, c(1,2) )

## of an sf object
sf <- sfheaders::sf_linestring(
  obj = df
  , x = "x"
  , y = "y"
)

calculate_bbox(sf$geometry, NULL )

```


```{r}

## Count Coordinates
##
## either counts the number of coordinates in a given geometry,
library(Rcpp)

cppFunction(
  depends = 'geometries'
  , includes = '#include "geometries/geometries/coordinates.hpp"'
  , code = '
    R_xlen_t count_coordinates( SEXP x ) {
      R_xlen_t count = 0;
      geometries::coordinates::count_coordinates( x, count );
      return count;
    }
  '
)

df <- data.frame(
  x = 1:10
  , y = 10:1
  , z = 21:30
  , m = 30:21
  , val = letters[1:10]
  , id = c( rep(1,5), rep(2,5) )
)

poly <- polygon( df, c("x","y"), c("id") )

count_coordinates( poly )
```

```{r}

## Coordinate indices
## 
## gives the start and end indices of geometries
cppFunction(
  depends = 'geometries'
  , includes = '#include "geometries/geometries/coordinates.hpp"'
  , code = '
    Rcpp::IntegerMatrix coordinate_indices( Rcpp::List x ) {
      return geometries::coordinates::coordinate_indices( x );
    }
  '
)

l <- list(
  matrix(1:20, ncol = 2 )
  , matrix(1:6, ncol = 2)
  , matrix(1:10, ncol = 2)
)

coordinate_indices( l )


```


```{r}

## id positions
## 
## takes a sorted vector and returns the start & end indices of each unique element
cppFunction(
  depends = 'geometries'
  , includes = '#include "geometries/utils/lines/lines.hpp"'
  , code = '
    Rcpp::IntegerMatrix id_positions( SEXP x ) {
      return geometries::utils::id_positions( x );
    }
  '
)

x <- c( rep(1, 5), rep(2, 10), rep(3, 6), rep(4, 2) )
id_positions( x )


```

